---
title: "Vegetation Departure for WDFFP Focal Areas"
author: "Randy Swaty"
---


<font size="4"> 
**These LANDFIRE-Powered Landscape Assessments allow users to start exploring trends, landscape conversion and natural resource management options. **

</font> 


## Goals of this demonstration

We will use [LANDFIRE](https://landfire.gov/){target="blank"} products to:

1. Characterize and map past and present ecosystem conditions
2. Explore historical disturbance patterns
3. Summarize patterns of change and management opportunities




## Location

The SRFSN Black Hills Subregion streches across portions of western South Dakota, eastern Wyoming and northwestern Nebraska.


<br>

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}

library(foreign)
library(sf)
library(terra)
library(tidyverse)
library(raster)


```


```{r read shapefile, message=FALSE, warning=FALSE, include=FALSE}
#  read shape
shp <- st_read("data/") %>% 
  st_transform(crs = 5070) %>%
  st_union() %>%
  st_sf()
```

```{r locator map, message=FALSE, warning=FALSE, echo=FALSE}
library(tmap)

# toggle tmap mode to interactive viewing
tmap_mode("view")

# create a quick interactive map
quickmap <- qtm(shp, 
                borders = "darkgreen", 
                fill = NULL, 
                check.and.fix = TRUE, 
                basemaps = 'https://server.arcgisonline.com/ArcGIS/rest/services/World_Topo_Map/MapServer/tile/{z}/{y}/{x}',
                title = 'Black Hills Subregion',)

quickmap



```

## How to use this web report

* All maps, input datasets and further assistance can be obtained by contacting the author, [Randy Swaty](mailto:rswaty@tnc.org){target="blank"}, Nature Conservancy Ecologist, Conservation Data Lab co-lead
* Review is ongoing
* To share this web app, simply share the hyperlink
* Toggle between dark and light display of this site in the upper right corner of the page
* Learn more about LANDFIRE and The Nature Conservancy [here](about.qmd){target="blank}.   



