---
title: "Wrangle and Calculate VDEP"
format: html
---

## Goal

## Packages

Aimed for a minimum; though went with the [Tidyverse](https://www.tidyverse.org/) way of coding and cheated column naming with the [janitor](https://sfirke.github.io/janitor/) package.

```{r}
#| label: load packages
#| echo: true
#| message: false
#| warning: false

library(janitor)
library(tidyverse)
```


## Input data

First data read in is the raw data from an ArcGIS [Combine](https://pro.arcgis.com/en/pro-app/3.3/tool-reference/spatial-analyst/combine.htm) of:

1. A rasterized file of the WDFFP focal areas, from the WDFFP "FSR_Firesheds_Projects_MEL.gdb" > FSR_PA.shp > non_null values from the "Focal_Geog" attribute.
2. LANDFIRE [Biophysical Settings (BpS)](https://landfire.gov/vegetation/bps)
3. LANDFIRE [Succession classes (Scls)](https://landfire.gov/vegetation/sclass) for 2020 and 2022

Minimal fields were joined in using the [Join Field](https://pro.arcgis.com/en/pro-app/latest/tool-reference/data-management/join-field.htm) tool in ArcGIS Pro.

Renaming of fields minimized; only changed when needed to prevent confusion.

Second input dataset is a wrangled "Reference Conditions" table, where the original (downloadable from https://landfire.gov/vegetation/bps-models) was modified in the following ways:

1. Current succession classes added: Agriculture, Developed, Water, UN and UE.
2. Dataset pivoted from wide to long (pivot_longer function from Tidyverse package)
3. model_label field was added by unioning the bps_model and label fields.  This was added for later use in joining. 
4. Fields rename using snake case.


```{r}
#| label: read in input datasets
#| echo: true
#| message: false
#| warning: false

base_data <- read.csv("data/combine.csv") %>%
  clean_names() %>%
  rename(c(
    scls_2020 = label,
    scls_2022 = label_1
  ))

ref_con <- read.csv("data/ref_con_long.csv")
```

## Calculate scls percents per pa/scls/bps combination.  

The key here is that we want the 2020 and 2022 scls percentages within each project area per scls for each bps_model.  Clear?

To do this we:

Remove unwanted sclss (labels are the same for 2020 and 2022-I checked.  Make sure to check moving forward).  Assuming unwanted scls values do not change over time.  This could introduce minor slop. 
Get a total count (of pixels) per PA
Get a total count for each BpS within each PA
Make sure that final calculations for 2020 and 2022 are different and make sense


```{r}
#| label: calculate scls percents
#| echo: true
#| message: false
#| warning: false

scls_labels <- setequal(unique(base_data$scls_2020), unique(base_data$scls_2022))
print(scls_labels) # should be true-so will use just one for removing unwanted rows

# add in total count per PA
data_current <- base_data %>%
  filter(!scls_2020 %in% c(
    "Barren or Sparse",
    "Water",
    "Snow/Ice",
    "Fill-Not Mapped"
  )) %>%
  group_by(wdffp_pas_r) %>%
mutate(total_count_pa = sum(count))

## matches exploration in Excel pivot table

# add in total count of BpS per PA
data_current <- data_current %>%
  group_by(wdffp_pas_r, bps_model) %>%
  mutate(total_count_pa_bps = sum(count))

## matches exploration in Excel pivot table

# add in count of each scls 2020 within each PA per BpS

data_current <- data_current %>%
  group_by(wdffp_pas_r, bps_model, scls_2020) %>%
  mutate(count_pa_bps_scls_2020 = sum(count))

## matches exploration in Excel pivot table

# add in count of each scls 2022 within each PA per BpS
      
data_current <- data_current %>%
  group_by(wdffp_pas_r, bps_model, scls_2022) %>%
  mutate(count_pa_bps_scls_2022 = sum(count))

## matches exploration in Excel pivot table


```







## Join in Reference Conditions

Below we create a new working dataframe, then add in reference percents for each bps_model - scls combination.  To do so we:

1. remove unwanted scls labels
2. create "model_label_2020" and model_label_2022" columns by using the union function
2. left_join ref_con model_labels for both model_label_2020 and model_label_2022 