---
title: "Wrangle and Calculate VDEP"
format: html
---

## Goal

## Packages

Aimed for a minimum; though went with the [Tidyverse](https://www.tidyverse.org/) way of coding and cheated column naming with the [janitor](https://sfirke.github.io/janitor/) package.

```{r}
#| label: load packages
#| echo: true
#| message: false
#| warning: false

library(janitor)
library(tidyverse)
```


## Input data


### Combine from ArcGIS pro for current conditions

First data read in is the raw data from an ArcGIS [Combine](https://pro.arcgis.com/en/pro-app/3.3/tool-reference/spatial-analyst/combine.htm) of:

1. A rasterized file of the WDFFP focal areas, from the WDFFP "FSR_Firesheds_Projects_MEL.gdb" > FSR_PA.shp > non_null values from the "Focal_Geog" attribute.
2. LANDFIRE [Biophysical Settings (BpS)](https://landfire.gov/vegetation/bps)
3. LANDFIRE [Succession classes (Scls)](https://landfire.gov/vegetation/sclass) for 2020 and 2022 separately
4. Exported attribute tables for [2020](data/combine2020.csv) and [2022](data/combine2022.csv)

Minimal fields were joined in using the [Join Field](https://pro.arcgis.com/en/pro-app/latest/tool-reference/data-management/join-field.htm) tool in ArcGIS Pro.

Renaming of fields minimized; only changed when needed to prevent confusion.

### Reference conditions

Second input dataset is a wrangled "Reference Conditions" table, where the original (downloadable from https://landfire.gov/vegetation/bps-models) was modified in the following ways:

1. Current succession classes added: Agriculture, Developed, Water, UN and UE.
2. Dataset pivoted from wide to long (pivot_longer function from Tidyverse package)
3. model_label field was added by unioning the bps_model and label fields.  This was added for later use in joining. 
4. Fields rename using snake case.

### Issues and futuring


### Reading in the datasets


```{r}
#| label: read in input datasets
#| echo: true
#| message: false
#| warning: false

base_data_2020 <- read.csv("data/combine2020.csv") %>%
  clean_names() 

base_data_2022 <- read.csv("data/combine2022.csv") %>%
  clean_names() 

ref_con <- read.csv("data/ref_con_long.csv")
```


## 2020 data wrangling and calculating

```{r}
#| label: calculate scls 2020 percents
#| echo: true
#| message: false
#| warning: false



# Add in total count per PA
data_2020 <- base_data_2020 %>%
  filter(!label %in% c(
    "Barren or Sparse",
    "Water",
    "Snow/Ice",
    "Fill-Not Mapped"
  )) %>%
  group_by(wdffp_pas_r) %>%
mutate(total_count_pa = sum(count))

## matches exploration in Excel pivot table

# Add in total count of BpS per PA
data_2020 <- data_2020 %>%
  group_by(wdffp_pas_r, bps_model) %>%
  mutate(total_count_pa_bps = sum(count))

## matches exploration in Excel pivot table

# Add in count of each scls 2020 within each PA per BpS

data_2020 <- data_2020 %>%
  group_by(wdffp_pas_r, bps_model, label) %>%
  mutate(count_pa_bps_scls_2020 = sum(count))


# Add in scls 2020 percents per BpS within each PA
data_2020 <- data_2020 %>%
  group_by(wdffp_pas_r, bps_model) %>%
  mutate(percent = 
           round((count_pa_bps_scls_2020/total_count_pa_bps) * 100, 1))

# Add a column for filtering the refcon (pa_bps_model)

# Add a column for joining 

data_2020 <- data_2020 %>%
    unite(pa_bps_scl, c("wdffp_pas_r", "bps_model", "label"), remove = FALSE)



write.csv(data_2020, file = 'data_2020.csv')
```

## 2022 data wrangling and calculating

```{r}
#| label: calculate scls 2022 percents
#| echo: true
#| message: false
#| warning: false



# Add in total count per PA
data_2022 <- base_data_2022 %>%
  filter(!label %in% c(
    "Barren or Sparse",
    "Water",
    "Snow/Ice",
    "Fill-Not Mapped"
  )) %>%
  group_by(wdffp_pas_r) %>%
mutate(total_count_pa = sum(count))

## matches exploration in Excel pivot table

# Add in total count of BpS per PA
data_2022 <- data_2022 %>%
  group_by(wdffp_pas_r, bps_model) %>%
  mutate(total_count_pa_bps = sum(count))

## matches exploration in Excel pivot table

# Add in count of each scls 2022 within each PA per BpS

data_2022 <- data_2022 %>%
  group_by(wdffp_pas_r, bps_model, label) %>%
  mutate(count_pa_bps_scls_2022 = sum(count))


# Add in scls 2022 percents per BpS within each PA
data_2022 <- data_2022 %>%
  group_by(wdffp_pas_r, bps_model) %>%
  mutate(percent = 
           round((count_pa_bps_scls_2022/total_count_pa_bps) * 100, 1))

# Add a column for joining 

data_2022 <- data_2022 %>%
    unite(pa_bps_scl, c("wdffp_pas_r", "bps_model", "label"), remove = FALSE)

write.csv(data_2022, file = 'data_2022.csv')
```



## Wrangle Ref Con 

Need a 'replicate' for each PA and addition of the PA number


```{r}
#| label: build refcon
#| echo: true
#| message: false
#| warning: false

# Get list of PAs
pas <- unique(base_data_2020$wdffp_pas_r)

# Filter ref_con to relevant BpSs

wdffp_bpss <- unique(base_data_2020$bps_model)

filtered_ref_con <- ref_con %>%
  filter(model_code %in% wdffp_bpss)


wdffp_ref_con <- do.call(rbind, lapply(pas, function(pas) {
  filtered_ref_con$pas <- pas
  return(filtered_ref_con)
}
  ))



```

